### 2016-09-27 analysis/paper/untitled.Rmd
add outline

---
title: 'The title'
author:
- Ben Marwick
- Maria Schaarschmidt
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   bookdown::word_document2:
    fig_caption: yes
    # reference_docx: templates/template.docx
bibliography: references.bib
# csl: journal-of-archaeological-science.csl
abstract: |
  This is the abstract.
---
```{r setup, include=FALSE, echo = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE)
opts_chunk$set(echo = FALSE, 
               messages = FALSE, 
               warnings = FALSE, 
               dpi = 300,
               cache = TRUE)
library(tidyverse)
```

# Introduction

# Background

## Geomorphology of the Irrawady terraces

- survey early literature on red earth deposits

## Archaeology of the Irrawady terraces

- Movius and contemporaries
- Dennell is a major commentator recently

## Survey data

```{r}

```

- our landscape observations
- descriptions of existing cuttings that we visited
- GPS data from transect
- surface finds of artefacts by KK

## Formation of the terraces and archaeological contexts

- summary of formation 
- implications for age of the archaeological deposit

## Conclusion

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

In Table \@ref(tab:carstab) we can see some data about the relationship between pressure and tempurature.

```{r carstab}
library(knitr)
kable(head(mtcars), caption = "Data about cars")
```

## Including Plots

You can also embed plots, for example:

```{r pressureplot, echo=FALSE, fig.cap="Plot of car data"}
plot(pressure)
```

In Figure \@ref(fig:pressureplot) we can see some data about pressure.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Citations and References

And we can have a citation, using a `.bib` file that holds all the details. To get this: [@Marwick2016repro] we type `[@Marwick2016repro]`. The text after the `@` is the bibtex key that links the in-text citation to the full details in the `.bib` file.

All of the usual variations on in-text citation formatting are possible in markdown, and listed for reference here: <http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html>

## Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
# which R packages and versions?
kable(data.frame(Setting = names(devtools::session_info()$platform),
                 Value = sapply(devtools::session_info()$platform, `[`, 1),
                 stringsAsFactors = FALSE,
                 row.names = NULL),
      caption = "R session information")
```

```{r}
kable(devtools::session_info()$packages,
      caption = "Packages that this report depends on")

# what commit is this file at?
library(git2r)
repo <- repository(path = "../..")
last_commit <- commits(repo)[[1]]
```

The current git commit of this file is `r last_commit@sha`, which is on the `r branches(repo)[[1]]@name` branch and was made by `r last_commit@committer@name` on `r when(last_commit)`. The current commit message is "`r last_commit@summary`".

## References