### 2016-10-04 analysis/paper/paper.Rmd
intro and early geo literature

---
title: 'The title'
author:
- Ben Marwick
- Maria Schaarschmidt
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   bookdown::word_document2:
    fig_caption: yes
    # reference_docx: templates/template.docx
bibliography: references.bib
# csl: journal-of-archaeological-science.csl
abstract: |
  This is the abstract.
---
```{r setup, include=FALSE, echo = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE)
opts_chunk$set(echo = FALSE, 
               messages = FALSE, 
               warnings = FALSE, 
               dpi = 300,
               cache = TRUE)
library(tidyverse)
```

# Introduction

The Irrawaddy River terraces of central Myanmar have been part of global Pleistocene archaeology since de Terra and Movius [-@deTerra1943] published their monograph reporting archaeological evidence from the Middle Pleistocene. Their expedition also resulted in Movius' [-@movius1948lower] influential paper claiming to identify a major cultural and technological division between the east and west hemispheres. Movius described a region to the west of the Irrawaddy where hand-axes are present in Middle Pleistocene contexts, and a region to the east where hand-axes are absent, with chopper-chopping tools found instead. This division has subsequently been known as the Movius Line, and has been the subject of extensive discussion by archaeologists [@lycett2010demographic; @petraglia2008large; @lycett2010movius; keates2002movius; @norton2006movius; @yamei2000mid; @norton2008movius; @brantingham1999astride; @brumm2012biface; @brumm2010movius; @norton2006middle; @davis1999recent; @norton2000current; @lycett2007soanian; Romanowska2016]. This includes critical discussion that has questioned the reliability of Movius' observations

The importance of the Movius' findings depend on the chronology of the artefacts he discovered.  inferred from his geomorphological interpretations. Despite the prominence of the Movius Line in discussions of global patterns of human evolution, there has not been any new data from the location of Line since Movius' initial field work. In this paper we assess the current status of the

# Background

The study area is located between the Sagaing fault to the west and the Shan scarp to the east. The Sagaing fault is a major strike-slip right-lateral continental fault that extends over 1200 km, first described by Noetling [-@noetling1900miocene], and later described by others [e.g. @thein1991lateral; maung1987transcurrent]. The Shan scarp is a topographic discontinuity that marks the boundary of the central plains and the Shan plateau to the east, a region with an average elevation of 1000 m and large variations of elevation over short distances (up to 1800 m over a few kilometers) [@bertrand2001diachronous; bertrand2003tectonics].

Descriptions of the physical and cultural geography of the Irrawaddy terraces of central Myanmar have been appearing scholarly literature since Myanmar was attached to the British Empire in 1886. In the following section we review this literature to characterise this landscape. Then we review previous work on the archaeology of this landscape. Finally, we report on our recent surveys in the Chauk region, and discuss the implications for the archaeological significance of the region.

## Geomorphology of the Irrawady terraces

Rep

chhibber1934geology

- survey early literature on red earth deposits

## Archaeology of the Irrawady terraces

- Movius and contemporaries
- Dennell is a major commentator recently

## Survey data

```{r}
# analysis of survey and GPS data in here
```

- our landscape observations
- descriptions of existing cuttings that we visited
- GPS data from transect
- surface finds of artefacts by KK

## Formation of the terraces and archaeological contexts

- summary of formation 
- implications for age of the archaeological deposit

## Conclusion

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

In Table \@ref(tab:carstab) we can see some data about the relationship between pressure and tempurature.

```{r carstab}
library(knitr)
kable(head(mtcars), caption = "Data about cars")
```

## Including Plots

You can also embed plots, for example:

```{r pressureplot, echo=FALSE, fig.cap="Plot of car data"}
plot(pressure)
```

In Figure \@ref(fig:pressureplot) we can see some data about pressure.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Citations and References

And we can have a citation, using a `.bib` file that holds all the details. To get this: [@Marwick2016repro] we type `[@Marwick2016repro]`. The text after the `@` is the bibtex key that links the in-text citation to the full details in the `.bib` file.

All of the usual variations on in-text citation formatting are possible in markdown, and listed for reference here: <http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html>

## Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
# which R packages and versions?
kable(data.frame(Setting = names(devtools::session_info()$platform),
                 Value = sapply(devtools::session_info()$platform, `[`, 1),
                 stringsAsFactors = FALSE,
                 row.names = NULL),
      caption = "R session information")
```

```{r}
kable(devtools::session_info()$packages,
      caption = "Packages that this report depends on")

# what commit is this file at?
library(git2r)
repo <- repository(path = "../..")
last_commit <- commits(repo)[[1]]
```

The current git commit of this file is `r last_commit@sha`, which is on the `r branches(repo)[[1]]@name` branch and was made by `r last_commit@committer@name` on `r when(last_commit)`. The current commit message is "`r last_commit@summary`".

## References