### 2016-10-06 analysis/paper/paper.Rmd
more on movius

---
title: 'The title'
author:
- Ben Marwick
- Maria Schaarschmidt
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   bookdown::word_document2:
    fig_caption: yes
    # reference_docx: templates/template.docx
bibliography: references.bib
# csl: journal-of-archaeological-science.csl
abstract: |
  This is the abstract.
---
```{r setup, include=FALSE, echo = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE)
opts_chunk$set(echo = FALSE, 
               messages = FALSE, 
               warnings = FALSE, 
               dpi = 300,
               cache = TRUE)
library(tidyverse)
```

# Introduction

The Irrawaddy River terraces of central Myanmar have been part of global Pleistocene archaeology since de Terra and Movius [-@deTerra1943] published their monograph reporting archaeological evidence from the Middle Pleistocene. Their expedition also resulted in Movius' [-@movius1948lower] influential paper claiming to identify a major cultural and technological division between the east and west hemispheres. Movius described a region to the west of the Irrawaddy where hand-axes are present in Middle Pleistocene contexts, and a region to the east where hand-axes are absent, with chopper-chopping tools found instead. This division has subsequently been known as the Movius Line, and has been the subject of extensive discussion by archaeologists [@lycett2010demographic; @petraglia2008large; @lycett2010movius; keates2002movius; @norton2006movius; @yamei2000mid; @norton2008movius; @brantingham1999astride; @brumm2012biface; @brumm2010movius; @norton2006middle; @davis1999recent; @norton2000current; @lycett2007soanian; Romanowska2016]. This includes critical discussion that has questioned the reliability of Movius' field observations [@dennell2016life; @dennell2014hallam].

The importance of the Movius' findings depend primarily on the chronology of the landscape he described, and the technology of the artefacts that he collected.  Despite the prominence of the Movius Line in discussions of global patterns of human evolution, there has not been any new data from the location of Line since Movius' initial field work. Movius' interpretation of the Irrawaddy terraces and the technology of the artefacts he described remains untested by new data from his study area. In this paper we assess the current status of the Irrawaddy terraces and stratigraphy, and the technology of the artefacts found near the Irrawaddy in central Myanmar. We report on our recent surveys and field observations of the key archaeological locations that Movius visited in January of 1938.

# Background

The study area is located between the Sagaing fault to the west and the Shan scarp to the east. The Sagaing fault is a major strike-slip right-lateral continental fault that extends over 1200 km, first described by Noetling [-@noetling1900miocene], and later described by others [e.g. @thein1991lateral; @maung1987transcurrent]. The Shan scarp is a topographic discontinuity that marks the boundary of the central plains and the Shan plateau to the east, a region with an average elevation of 1000 m and large variations of elevation over short distances (up to 1800 m over a few kilometers) [@bertrand2001diachronous; @bertrand2003tectonics].

Descriptions of the physical and cultural geography of the Irrawaddy terraces of central Myanmar began appearing with increasing frequency in scholarly literature after Myanmar was attached to the British Empire in 1886. The Japanese occupation in 1942 disrupted geological research, and few publications appeared after the Second World War [@bender1983geology].  In the following sections we review this literature to characterise this landscape. Then we review previous work on the archaeology of this landscape.

## Geomorphology of the Irrawady terraces

Rep

chhibber1934geology

- survey early literature on red earth deposits

## Archaeology of the Irrawady terraces

Since the 19th century, the focus of archaeological interest lay on the central part of the Irrawaddy River in central Myanmar. The first artefacts were discovered by Morris in 1930 in Upper Myanmar and later assigned to the “new” Anyathian culture from the Lower Palaeolithic in Southeast Asia. The Anyathian artefacts are mainly not found in situ but in the secondary deposits of the Pleistocene terrace gravels of the Irrawaddy River. The occurrence of only few handaxes is one main characteristic of the Anyanthian culture. The occupation of the Irrawaddy River terraces during the Pleistocene seems to be limited to the east bank of the River and the central part of Myanmar because the finds concentrate in the region around Chauk, Nyaung-U and Yenanyaung [@movius1944early; @deTerra1943].

The three phases of the Early Anyathian are all dominated by a high degree of uniformity of the stone tools, probably due to raw material constraints (fossil wood and silicified tuff). Fossil wood is difficult to flake because it easily flakes parallel to the natural fibres of the wood. Most of the artefacts made from fossil wood are worked on only one plane; multiple flake scars are rare. Silicified tuff is more suitable for stone tool production due to its fine-grained homogenous texture and conchodial fractures, although it can be very porous. In total, 483 implements were found (261 from fossil wood, 220 from silicified tuff and 2 from quartzite). Implements made from fossil wood display a limited range of tools, including mostly handaxes, as well as choppers and chopping tools. Generally, stone tools made of silicified tuff are more massive than tools made from fossil wood. Choppers are dominant, whereas there are few handaxes [@deTerra1943].

In addition, 23 early Anyathian artefacts were found in situ in the lateritic gravel and gravel deposits of terrace 1 near Chauk on the Hill of Chinaungma. Movius (1943) assigned these implements to the first phase of the Early Anyathian and the beginning of the Middle Pleistocene as the oldest human artefacts found in the Irrawaddy Valley. Implements of phase 2 of the Early Anyathian show either a ferruginous crust, or are heavily rolled. The former implies an occupation of this area during the forming of the crust in a long interpluvial period. The latter suggests that these implements were derived from older deposits. Most of the Artefacts from the third phase of the Early Anyathian are very heavily rolled and are associated with the early Upper Pleistocene when the gravels were deposited [@deTerra1943].

Norton et al. [-@norton2006middle] reworked the concept of the Movius Line by incorporating existing issues with the “traditional” Movius Line. The three main characteristics of the “Movius Line sensu lato” are the lower frequency of handaxe bearing sites in East Asia compared to Africa and India; the much lower percentage of bifacially made tools in East Asian assemblages and the morphological similarities to Acheulean artefacts [@norton2006middle; @norton2008movius].

Instead of assigning stone tools to certain technologies made by different groups of early humans, and making implications about their abilities to produce stone tools, acknowledging the spatial and temporal diversity of lithic records in Eurasia is necessary [@dennell2016life]. Reasons for the lack of Acheulean technology are investigated in modern quaternary science [@brumm2012biface; @dennell2016life; @norton2008movius; @norton2006middle; @petraglia2008large; @schick1994movius], including constraints on raw material, demographic and social transmission, environmental changes and dispersal routes.

The use of stone tools may have been influenced by certain situations, tasks or individuals. Consequently, different members of a group or different groups could have used bifaces rarely, never or only at specific events or times. Thus, the absence of bifaces or an item in general does not mean that the hominin living in this area at this time was not able to produce such a tool [@dennell2016life].

Movius [-@movius1944early] stated that the raw material used east of the Movius Line was often of low quality quartz and quartzite that would prevent the production of the same kind of bifaces [@dennell2016life; @lycett2010movius]. However, handaxes from Zhoukoudian Locality 1, Chongokni and Kumpari (China) are bifacially worked and are made of quartz and quartzite river cobbles from this area [@lycett2010movius].

Whilst colonizing East Asia, early modern humans encountered barriers such as mountain ranges, river deltas, oceans and deserts that provide a range of resources (e.g. water, food, shelter) [@lycett2010movius]. However, the environment during the Middle Pleistocene was challenging in terms of changing climatic conditions and accompanying biogeographic transitions [@bar2001africa; @field2005assessment; @lycett2010movius; @schick1994movius].

There may have been no need for intensive stone tool production, or a different material was used at some point. This theory is known as the “bamboo hypothesis” [@field2005assessment; @lycett2010movius; @schick1994movius] and states that early modern humans used bamboo instead of stones to make tools [@schick1994movius] leaving behind no archaeological record [@lycett2010movius]. However, even though bamboo cutting marks on bones can be distinguished from other cutting marks, there have been no findings of such so far [@lycett2010movius].

The dispersal route of early modern humans from Africa to Asia and Australia is still debated among quaternary scientists. Based on GIS analyses, Field & Lahr [-@field2005assessment] and Field et al. [-@field2007southern] identified possible routes from Africa leading eastwards along the coasts to Asia and eventually via the Sunda Shelf to Australia. This “Southern Dispersal Route” is dated to around 75-60ka [@field2005assessment; @field2007southern; @clarkson2012continuity; @macaulay2005single]. This route matches with some Middle Pleistocene archaeological sites in Africa, India (e.g. Borra, the Kokan Complex and the Hiran Valley) and SE Asia adjacent to coastal regions. Of particular importance are the deposits at the Hiran Valley since they have been dated to 56-69ka, matching the proposed time for the Southern Dispersal. A dispersal along the coasts and rivers seems reasonable because they provide sufficient resources. However, these first temporary settlements would not necessarily leave behind an archaeological record due to the rising sea level and the subsequent flooding of the Sunda shelf because of environmental changes QUELLE.

Recently, the model of demographic and social transmission has been proposed as a possible explanation for the diverse development of technologies in E/SE Asia and W/SW Asia. The concept is based on the assumption that a certain effective population size is essential for developing traditions and technologies to be passed on to further generations [@lycett2010movius; @lycett2010demographic; @lycett2007there]. Given the colonization of the whole of East Asia and the distance to Africa decreasing the population size, the population density may have been too low to maintain or establish more elaborate tool making techniques [@lycett2010demographic; @lycett2007there].

Acheulean bifacial tools are the earliest known artefacts from Africa dating to 1.76 Ma respectively to 1.7- 1.6 Ma [@brumm2012biface; @lycett2010movius; @norton2006middle]. In contrary to Mode I technology, these stone tools show a degree of standardization [@norton2006middle] and were the main tools used by hominins during the Pleistocene [@petraglia2008large]. The tools include handaxes, cleavers, picks, knives, lanceolates and unifaces [@brumm2012biface]. Used until 100 ka ago, Acheulean tools are commonly interpreted as butchery tools, although other functions are possible as well [@brumm2012biface]. The definition of the Acheulean is not always straightforward, i.e. the assignment of finds can be influenced by the finding location east or west of the Movius Line QUELLE. Also, bifaces are not suitable as a time marker for specific cultures as they have been produced in different parts of the world over several 100 ka [@dennell2016life].

Dennell is one of the more recent critics of the concept of the Movius Line. In his analysis of Movius’ publications, he describes their views as “backwards”, “ancient” and “eurocentric” [@dennell2016life]. The drawing of a line to mark difference in stone tool production between SE and SW/W Asia neglects the variety and complexity of lithic assemblages on either side of the line, leaving East Asia in a minor position in human evolution. According to Dennell, none of the material found by Movius and his colleagues has a stratigraphic context and they failed to identify a sequence of four terraces along the Irrawaddy River from the Middle to Upper Pleistocene. Furthermore, because of the finding context, his finds cannot be connected to other Middle Pleistocene Acheulean assemblages in Southwest Asia [@dennell2016life].

## Survey data

```{r}
# analysis of survey and GPS data in here
```

- our landscape observations
- descriptions of existing cuttings that we visited
- GPS data from transect
- surface finds of artefacts by KK

## Formation of the terraces and archaeological contexts

- summary of formation 
- implications for age of the archaeological deposit

## Conclusion

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

In Table \@ref(tab:carstab) we can see some data about the relationship between pressure and tempurature.

```{r carstab}
library(knitr)
kable(head(mtcars), caption = "Data about cars")
```

## Including Plots

You can also embed plots, for example:

```{r pressureplot, echo=FALSE, fig.cap="Plot of car data"}
plot(pressure)
```

In Figure \@ref(fig:pressureplot) we can see some data about pressure.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Citations and References

And we can have a citation, using a `.bib` file that holds all the details. To get this: [@Marwick2016repro] we type `[@Marwick2016repro]`. The text after the `@` is the bibtex key that links the in-text citation to the full details in the `.bib` file.

All of the usual variations on in-text citation formatting are possible in markdown, and listed for reference here: <http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html>

## Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
# which R packages and versions?
kable(data.frame(Setting = names(devtools::session_info()$platform),
                 Value = sapply(devtools::session_info()$platform, `[`, 1),
                 stringsAsFactors = FALSE,
                 row.names = NULL),
      caption = "R session information")
```

```{r}
kable(devtools::session_info()$packages,
      caption = "Packages that this report depends on")

# what commit is this file at?
library(git2r)
repo <- repository(path = "../..")
last_commit <- commits(repo)[[1]]
```

## References